---
title: "Analysis of Climate Warming"
author: "Ruiyi Lian"
date: "2020/9/26"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE,
                      warning = FALSE,
                      message=FALSE)
library('dplyr')
library('reshape2')
library('forcats')
library('ggplot2')
library('stringr')
library('data.table')
library('lubridate')
library('zoo')
Sys.setenv(LANGUAGE="en")
library(forecast)
library(seastests)
library('Kendall')
setwd("D:/R_l7/data/data")
```

# Load and Clean Data  

Assemble the data into a single data frame for analysis.  
```{r}
dt <- data.table()
fileList <- list.files("D:/R_l7/data/data")
for (fileName in fileList) {
  setwd("D:/R_l7/data/data")
  dtSingle = fread(fileName)
  dt <- rbind(dt,dtSingle,fill=TRUE)
}
head(dt)
```

Transform the date-time data into posix numbers using lubridate and make substitutions for NA data.  

```{r}
dt$Year <- paste0(dt$YYYY,dt$`#YY`)
dt$Year <- str_replace(dt$Year,"NA","")
dt$date <- as.Date(paste0(dt$Year,"-",dt$MM,"-",dt$DD))
# Transform the date-time data into posix numbers using lubridate
dt$date <- as_datetime(dt$date)
```

Make a Backup file.  

```{r}
dtBackup <- dt
```

Determine month frequency.  

```{r}
dt <- dtBackup[,c("date","ATMP","WTMP")]
dt[,":="(ATMP=as.numeric(ATMP),
         WTMP=as.numeric(WTMP))]
dt$ATMP[dt$ATMP==999]=NA
dt$WTMP[dt$WTMP==999]=NA
dt$ym <- as.yearmon(dt$date)
dt <- na.omit(dt)
dtMonth <- dt[,.(ATMP=mean(ATMP,na.rm = T),
            WTMP=mean(WTMP,na.rm = T)),by=ym]
length(unique(dtMonth$ym))
names(dtMonth) <- c("ym","Air","Water")
dtMonth <- na.omit(dtMonth)
```

# Visualization  

Compare temperature change in air vs water.  

```{r}
dtMP <- melt(dtMonth,id.vars = "ym")
ggplot(dtMP,aes(ym,value,color=variable))+
  geom_line()+
  labs(x="",y="Temperature (°F)")+
  theme(legend.title = element_blank())
```

Seasonal of air temperature.  

```{r}
AirTemp <- ts(c(dtMonth$Air[c(1:156)],
                rep(NA,6),
                dtMonth$Air[c(157:247)]),
              start = 1999,
              frequency = 12)
```


```{r}
ggseasonplot(AirTemp)
```

Seasonal of water temperature.  

```{r}
WaterTemp <- ts(c(dtMonth$Water[c(1:156)],
                rep(NA,6),
                dtMonth$Water[c(157:247)]),
              start = 1999,
              frequency = 12)
ggseasonplot(WaterTemp)
```


```{r}
WaterTemp <- ts(c(dtMonth$Water[c(1:156)],
                rep(NA,6),
                dtMonth$Water[c(157:247)]),
              start = 1999,
              frequency = 12)
ggseasonplot(WaterTemp)
```

# test  

Before use Mann-Kendall trend test we need check whether there is seasonal in data or not.  

```{r}
isSeasonal(WaterTemp, freq = 12)
```

The result indicated that there is seasonal in data.  
So  we use Seasonal Mann-Kendall trend test to a test for monotonic trend in a time series.  

```{r}
SeasonalMannKendall(WaterTemp)
SeasonalMannKendall(AirTemp)
```

# Conclusion  

As p-value is smaller than 0.05, we reject H0 and determine that there are monotonic trend in both air and water temperature from 1999 to 2009.  